---
title: "ACT03 - Data Mining"
author: "Team 7 | Captain: Nadine Rose | Members: Kyle Scott, Lakshya Rathore, Bailey LaRea"
date: \today
documentclass: report
geometry: margin=2cm
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage{mathtools}
    - \usepackage{physics}
    - \usepackage{enumitem}
    - \usepackage{color}
    - \usepackage{soul}
    - \usepackage{graphicx}
output: pdf_document
---

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1.1

```{r, warning=FALSE, message=FALSE}
library(kableExtra)

Observation <- c(1:12)
True_Status <- c(1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0)
Posterior_Probability <- c(0.95, 0.85, 0.75, 0.45, 0.35, 0.25, 0.15, 0.05, 0.65, 0.55, 0.5, 0.7)

df <- data.frame(Observation, True_Status, Posterior_Probability)

kable(df, linesep = "\\addlinespace", 
      digits = 3, booktabs = T, format = 'latex') %>% 
  kable_styling(latex_options = c("striped", "hold_position"))
```

We are given that the cut-off point is 0.72. Using this chart, we know that any Posterior Probability $\geq 0.72$ will be classified as 1 and any Posterior Probability $< 0.72$ will be classified as 0.

```{r}
Predicted_Status <- c(1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0)
df <- cbind(df, Predicted_Status)

kable(df, linesep = "\\addlinespace", 
      digits = 3, booktabs = T, format = 'latex') %>% 
  kable_styling(latex_options = c("striped", "hold_position"))
```

Now we can compare True_Status to Predicted_Status to determine the number of True Positives, False Positives, True Negatives and False Negatives.

### Answer Section:

True Positive (True\_Status = 1 \&\& Predicted\_Status = 1) = 3 (Obs. 1, 2, and 3)

False Positive (True\_Status = 0 \&\& Predicted\_Status = 1) = 0

True Negative (True\_Status = 0 \&\& Predicted\_Status = 0) = 6 (Obs. 7, 8, 9, 10, 11, and 12)

False Negative (True\_Status = 1 \&\& Predicted\_Status = 0) = 3 (Obs. 4, 5, and 6)

Sensitivity = $\frac{N_{TP}}{N_{TP} + N_{FN}} = \frac{3}{3 + 3} = 0.5$

Specificity = $\frac{N_{TN}}{N_{TN} + N_{FP}} = \frac{6}{6 + 0} = 1$

Accuracy = $\frac{N_{TP} + N_{TN}}{N} = \frac{3 + 6}{12} = 0.75$

Precision = $\frac{N_TP}{N_{TP} + N_{FP}} = \frac{3}{3 + 0} = 1$

F1 Score = $\frac{2\times(\text{Precision}\times\text{Sensitivity})}{\text{Precision} + \text{Sensitivity}} = \frac{2\times1\times0.5}{0.5 + 1} = 0.667$

# Problem 1.2

```{r, warning=FALSE, message=FALSE}
Rank <- c(12, 11, 10, 5, 4, 3, 2, 1, 8, 7, 6, 9)
Rank_Positive <- c(12, 11, 10, 5, 4, 3, 0, 0, 0, 0, 0, 0)
df <- cbind(df, Rank)
df <- cbind(df, Rank_Positive)
kable(df, linesep = "\\addlinespace", 
      digits = 3, booktabs = T, format = 'latex') %>% 
  kable_styling(latex_options = c("striped", "hold_position"))
```

```{r, warning=FALSE, message=FALSE}
sum_rankpos <- sum(Rank_Positive)
sum_rankpos
```

AUC = $\frac{\text{sumrankpos - 0.5} \times \pi N \times (\pi N + 1)}{\pi N (N - \pi N)} = \frac{45 - 0.5 \times 0.5 \times 12 \times (0.5 \times 12 + 1)}{0.5 \times 12 \times(12 - 0.5 \times 12)} = 0.667$

# Problem 1.3

GINI = $2 \times (\text{AUC} - 0.5) = 2 \times (0.667 - 0.5) = 0.3334$

# Problem 2

### Calculate KS Statistics

```{r, warning=FALSE, message=FALSE}
# create columns for Decile, Positive, and Negative as given in the Assignment
Decile <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
Positive <- c(100, 98, 96, 90, 85, 80, 75, 66, 51, 41)
Negative <- c(0, 2, 4, 10, 15, 20, 25, 34, 49, 59)

# create a data frame with these 3 columns
ks <- data.frame(Decile, Positive, Negative)
```

```{r, warning=FALSE, message=FALSE}
# create vars for Total Num of Positives and Negatives
total_pos <- sum(Positive)
total_neg <- sum(Negative)

# calculate TPR
TPR <- Positive / total_pos

# calculate TNR
TNR <- Negative / total_neg

# calculate cumulative TPR
cumulative_TPR <- cumsum(TPR)

# calculate cumulative TNR
cumulative_TNR <- cumsum(TNR)

# calculate KS statistic
KS <- cumulative_TPR - cumulative_TNR
```

```{r, warning=FALSE, message=FALSE}
# bind TPR to the data frame
ks <- cbind(ks, TPR)

# bind TNR to the data frame
ks <- cbind(ks, TNR)

# bind cumulative_TPR to the data frame
ks <- cbind(ks, cumulative_TPR)

# bind cumulative_TNR to the data frame
ks <- cbind(ks, cumulative_TNR)

# bind KS to the data frame
ks <- cbind(ks, KS)
```

```{r, warning=FALSE, message=FALSE}
library(kableExtra)

kable(ks, caption = "Calculating KS Statistics", linesep = "\\addlinespace", 
      digits = 3, booktabs = T, format = 'latex') %>% 
  kable_styling(latex_options = c("striped", "hold_position"))
```


# Problem 3.1

### Read in the CSV file for Microsoft

```{r, warning=FALSE, message=FALSE}
library(readr)

microsoft <- read.csv("Microsoft_Results.csv")
```

# Problem 3.2

### First Step is to create a function that will calculate all the required values which can be passed to the function parameter called 'cutoff'

```{r}
stats <- function(cutoff){

predicted_detections <- ifelse(microsoft$P_HasDetections >= cutoff, 1, ifelse(microsoft$P_HasDetections < cutoff, 0, NA))

TP <- sum(ifelse((microsoft$HasDetections == 1 & predicted_detections == 1), 1, 0))

TruePos <<- c(TruePos, TP)

FP <- sum(ifelse((microsoft$HasDetections == 0 & predicted_detections == 1), 1, 0))

FalsePos <<- c(FalsePos, FP)

TN <- sum(ifelse((microsoft$HasDetections == 0 & predicted_detections == 0), 1, 0))

TrueNeg <<- c(TrueNeg, TN)

FN <- sum(ifelse((microsoft$HasDetections == 1 & predicted_detections == 0), 1, 0))

FalseNeg <<- c(FalseNeg, FN)

Sensitivity <<- c(Sensitivity, TP / (TP + FN))

Specificity <<- c(Specificity, TN / (TN + FP))

Accuracy <<- c(Accuracy, (TP + TN) / (TP + TN + FP + FN))

Precision <<- c(Precision, TP / (TP + FP))
}
```

### Now that the function is created, we can call the function for whichever desired values we have. For the purpose of simplicity and demonstration, here we have chosen to display a table with all the desired statistics at the five percent level

```{r, results='hide'}
TP <- vector()
FP <- vector()
TN <- vector()
FN <- vector()
TruePos <- vector()
FalsePos <- vector()
TrueNeg <- vector()
FalseNeg <- vector()
Sensitivity <- vector()
Specificity <- vector()
Accuracy <- vector()
Precision <- vector()

cutoff <- seq(0.05, 1, by = 0.05)
mapply(stats, cutoff)
```

```{r}
stats_df <- data.frame(cutoff, TruePos, FalsePos, TrueNeg, FalseNeg, Sensitivity, Specificity, Accuracy, Precision)

kable(stats_df, caption = "Statistics for the five percentile level", linesep = "\\addlinespace", 
      digits = 3, booktabs = T, format = 'latex') %>% 
  kable_styling(latex_options = c("striped", "hold_position"))
```

# Problem 3.3

```{r, warning=FALSE, message=FALSE}
library(pROC)

AUC <- auc(microsoft$HasDetections, microsoft$P_HasDetections)
AUC
```

```{r, warning=FALSE, message=FALSE}
GINI <- 2*(AUC -  0.5)
GINI
```

# Problem 3.4

### Call our function from 3.2 but this time only for deciles

```{r, results = 'hide'}
TP <- vector()
FP <- vector()
TN <- vector()
FN <- vector()
TruePos <- vector()
FalsePos <- vector()
TrueNeg <- vector()
FalseNeg <- vector()
Sensitivity <- vector()
Specificity <- vector()
Accuracy <- vector()
Precision <- vector()

cutoff <- seq(0.1, 1, by = 0.1)
mapply(stats, cutoff)
```

### Create the ROC plot by graphing 1 - Specificity vs Sensitivity with some of the data provided by our function

```{r, warning=FALSE}
library(ggplot2)
Specificity <- 1 - Specificity
df <- data.frame(Specificity, Sensitivity)
plt <- ggplot(data = df, aes(x = Specificity, y = Sensitivity)) +
  geom_point(color = '#0072B2') + 
  geom_line(alpha = 0.3, color = '#0072B2') +
  geom_abline(slope = 1, inctercept = 0, color = '#0072B2') +
  ggtitle(label = "ROC Curve for Deciles (Ten Percent Level)") +
  xlab("1 - Specificity") +
  ylab("Sensitivity") +
  theme_light()

plt
```

# Problem 3.5

### Same process as 3.4 except this time we are working at the fifth percentile level

```{r, results = 'hide'}
TP <- vector()
FP <- vector()
TN <- vector()
FN <- vector()
TruePos <- vector()
FalsePos <- vector()
TrueNeg <- vector()
FalseNeg <- vector()
Sensitivity <- vector()
Specificity <- vector()
Accuracy <- vector()
Precision <- vector()

cutoff <- seq(0.05, 1, by = 0.05)
mapply(stats, cutoff)
```

```{r, warning=FALSE}
library(ggplot2)
Specificity <- 1 - Specificity
df <- data.frame(Specificity, Sensitivity)
plt <- ggplot(data = df, aes(x = Specificity, y = Sensitivity)) +
  geom_point(color = '#0072B2') + 
  geom_line(alpha = 0.3, color = '#0072B2') +
  geom_abline(slope = 1, inctercept = 0, color = '#0072B2') +
  ggtitle(label = "ROC Curve for the Five Percent Level") +
  xlab("1 - Specificity") +
  ylab("Sensitivity") +
  theme_light()

plt
```

# Problem 3.6

```{r}
true_pos <- function(decile){
  
vec <- microsoft$HasDetections[microsoft$P_HasDetections > decile - 0.1 & microsoft$P_HasDetections <= decile]

n <- length(vec)
  
count <- sum(vec == 1)
  
TPC <<- c(TPC, count)

N <<- c(N, n)
}
```

```{r, results='hide'}
total_pos <- sum(microsoft$HasDetections == 1)
prop_pos <- total_pos / 1000000

vec <- vector()
n <- vector()
count <- vector()
TPC <- vector()
N <- vector()

decile <- seq(1, 0.1, -0.1)

mapply(true_pos, decile)
```

```{r, warning=FALSE}
library(ggplot2)
lift <- (TPC / N) / prop_pos
new_seq <- seq(0.1, 1, 0.1)
lift_cum <- (cumsum(TPC) / cumsum(N)) / prop_pos
df <- data.frame(lift, lift_cum, new_seq)
plt <- ggplot(data = df, aes(x = new_seq)) +
  geom_point(aes(y = lift), color = '#0072B2') + 
  geom_line(aes(y = lift, color = "Lift"), alpha = 0.3) +
  geom_point(aes(y = lift_cum), color = '#CC79A7') + 
  geom_line(aes(y = lift_cum, color = "Cumulative Lift"), alpha = 0.3) +
  geom_hline(yintercept = 1, alpha = 0.5) +
  scale_color_manual("", breaks = c("Lift", "Cumulative Lift"), values = c('#0072B2', '#CC79A7')) +
  ggtitle(label = "Lift Chart") +
  xlab("Fraction") +
  ylab("Lift") +
  scale_x_continuous(n.breaks = 10) +
  theme_light()

plt
```

# Chapter IV Exercises

### Problem 1
\begin{figure}
\fbox{\includegraphics[width=11cm, height=22cm]{q1.png}}
\end{figure}

### Problem 9

\begin{enumerate}[label = (\alph*)]
\item \begin{enumerate}[label = (\roman*)] \item $\frac{p(X)}{1 - p(X)} = 0.37$ \item $p(X) = 0.37(1 - p(X))$ \item $1.37 \times p(X) = 0.37$ \item $p(X) = 0.37/1.37 =$ \hl{27\%} \end{enumerate}
\item Odds = $\frac{p(X)}{1 - p(X)} = 0.16/0.84 =$ \hl{0.19}
\end{enumerate}