---
title: "ACT05 - Data Mining"
author: "Team 7 | Captain: Nadine Rose | Members: Kyle Scott, Lakshya Rathore, Bailey LaRea"
date: \today
documentclass: report
geometry: margin=2cm
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage{mathtools}
    - \usepackage{physics}
    - \usepackage{enumitem}
    - \usepackage{color}
    - \usepackage{soul}
    - \usepackage{graphicx}
    - \usepackage{float}
output: pdf_document
---

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\section{Problem 1 - Programming}

\subsection{One - Read Data}

```{r, warning=FALSE, message=FALSE}
# read in the csv
library(readr)
hd <- read_csv("HeartDisease.csv")
```

\subsection{Two - Univariate Statistics}

```{r, warning=FALSE, message=FALSE, fig.pos='H'}
library(psych)
library(kableExtra)
library(dplyr)

# select only the needed columns and gather statistics
df <- select(hd, -c(names, famhist, chd))
tbl <- describe(df, skew = TRUE)
# calculate vairance
variance <- (tbl$sd)^2
variance <- round(variance, digits = 3)
# calculate first quartile
first_quartile <- summarize_all(df, ~ quantile(.x, 0.25))
first_quartile <- as.numeric(first_quartile[1,])
# calculate third quartile
third_quartile <- summarize_all(df, ~quantile(.x, 0.75))
third_quartile <- as.numeric(third_quartile[1,])
# bind extra statistics to existing tbl
tbl <- cbind(tbl, variance, first_quartile, third_quartile)
# reorder tbl
tbl <- tbl[, -c(1, 6, 7)]
tbl <- tbl[, c(2, 4, 8, 9, 3, 11, 10, 12, 13)]
# print tbl
kable(tbl, caption = "Univariate Statistics for Heart Disease Data", linesep = "\\addlinespace", digits = 3, 
      booktabs = T, format = 'pandoc')
```

\subsection{Three - Histograms}

```{r, warning=FALSE, message=FALSE, fig.show="hold", out.width="50%"}
library(ggplot2)
library(ggthemes)

plt_sbp <- ggplot(data = hd, aes(x = sbp)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$sbp),
                                         sd = sd(hd$sbp)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Systolic Blood Pressure") +
  xlab("Systolic Blood Pressure (sbp)") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_sbp

plt_tobacco <- ggplot(data = hd, aes(x = tobacco)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$tobacco),
                                         sd = sd(hd$tobacco)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Tobacco") +
  xlab("Tobacco") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_tobacco

plt_ldl <- ggplot(data = hd, aes(x = ldl)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$ldl),
                                         sd = sd(hd$ldl)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Low Density Lipoprotein") +
  xlab("Low Density Lipoprotein (ldl)") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_ldl

plt_adiposity <- ggplot(data = hd, aes(x = adiposity)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$adiposity),
                                         sd = sd(hd$adiposity)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Adiposity") +
  xlab("Adiposity") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_adiposity

plt_typea <- ggplot(data = hd, aes(x = typea)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$typea),
                                         sd = sd(hd$typea)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Type A Behavior") +
  xlab("Typea") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_typea

plt_obesity <- ggplot(data = hd, aes(x = obesity)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$obesity),
                                         sd = sd(hd$obesity)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Obesity") +
  xlab("Obesity") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_obesity

plt_alcohol <- ggplot(data = hd, aes(x = alcohol)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$alcohol),
                                         sd = sd(hd$alcohol)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Current Alcohol Consumption") +
  xlab("Alcohol") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_alcohol

plt_age <- ggplot(data = hd, aes(x = age)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$age),
                                         sd = sd(hd$age)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Age at Onset") +
  xlab("Age") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_age
```

\subsection{Four - Quantile Plots}

```{r, warning=FALSE, message=FALSE, fig.show="hold", out.width="50%"}
qplt_sdp <- ggplot(data = hd, aes(sample = sbp)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - sdp") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_sdp

qplt_tobacco <- ggplot(data = hd, aes(sample = tobacco)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - Tobacco") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_tobacco

qplt_ldl <- ggplot(data = hd, aes(sample = ldl)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - ldl") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_ldl

qplt_adiposity <- ggplot(data = hd, aes(sample = adiposity)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - adiposity") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_adiposity

qplt_typea <- ggplot(data = hd, aes(sample = typea)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - typea") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_typea

qplt_obesity <- ggplot(data = hd, aes(sample = obesity)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - obesity") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_obesity

qplt_alcohol <- ggplot(data = hd, aes(sample = alcohol)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - alcohol") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_alcohol

qplt_age <- ggplot(data = hd, aes(sample = age)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - age") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_age
```

\subsection{Five - Logistic Regression Model}

```{r}
logit_model <- glm(formula = chd ~ . - names - famhist, family = binomial(link = "logit"), data = hd)

summary(logit_model)
```

\subsection{Six - Power Transformations}

```{r, warning=FALSE, message=FALSE}
library(forecast)
pwr_sbp <- (hd$sbp) ^ (-2)
pwr_tobacco <- (hd$tobacco) ^ (0.4)
pwr_ldl <- (hd$ldl) ^ (0.1)
pwr_obesity <- (hd$obesity) ^ (-0.4)
pwr_alcohol <- (hd$alcohol) ^ (0.4)
```

```{r}
hd <- cbind(hd, pwr_sbp, pwr_tobacco, pwr_ldl, pwr_obesity, pwr_alcohol)
```

\subsection{Seven - Power Transformation Histograms}

```{r, warning=FALSE, message=FALSE, fig.show="hold", out.width="50%"}
library(ggplot2)
library(ggthemes)

plt_pwrsbp <- ggplot(data = hd, aes(x = pwr_sbp)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$pwr_sbp),
                                         sd = sd(hd$pwr_sbp)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Systolic Blood Pressure - Power Transformation (power = -2)") +
  xlab("Systolic Blood Pressure (sbp)") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_pwrsbp

plt_pwrtobacco <- ggplot(data = hd, aes(x = pwr_tobacco)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$pwr_tobacco),
                                         sd = sd(hd$pwr_tobacco)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Tobacco - Power Transformation (power = 0.4)") +
  xlab("Tobacco") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_pwrtobacco

plt_pwrldl <- ggplot(data = hd, aes(x = pwr_ldl)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$pwr_ldl),
                                         sd = sd(hd$pwr_ldl)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Low Density Lipoprotein (ldl) - Power Transformation (power = 0.1)") +
  xlab("Low Density Lipoprotein (ldl)") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_pwrldl

plt_pwrobesity <- ggplot(data = hd, aes(x = pwr_obesity)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$pwr_obesity),
                                         sd = sd(hd$pwr_obesity)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Obesity - Power Transformation (power = -0.4)") +
  xlab("Obesity") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_pwrobesity

plt_pwralcohol <- ggplot(data = hd, aes(x = pwr_alcohol)) + 
  geom_histogram(aes(y = ..density..), col = "black", fill = "#E69F00", bins = 15) +
  stat_function(fun = dnorm, args = list(mean = mean(hd$pwr_alcohol),
                                         sd = sd(hd$pwr_alcohol)), 
                col = "#009E73", size = 2) +
  ggtitle(label = "Alcohol - Power Transformation (power = 0.4)") +
  xlab("Alcohol") +
  ylab("Density") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))
  
plt_pwralcohol
```

\subsection{Eight - Power Transformation Quantile Plots}

```{r, warning=FALSE, message=FALSE, fig.show="hold", out.width="50%"}
qplt_pwrsdp <- ggplot(data = hd, aes(sample = pwr_sbp)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - sdp (power = -2)") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_pwrsdp

qplt_pwrtobacco <- ggplot(data = hd, aes(sample = pwr_tobacco)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - Tobacco (power = 0.4)") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_pwrtobacco

qplt_pwrldl <- ggplot(data = hd, aes(sample = pwr_ldl)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - ldl (power = 0.1)") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_pwrldl

qplt_pwrobesity <- ggplot(data = hd, aes(sample = pwr_obesity)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - obesity (power = -0.4)") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_pwrobesity

qplt_pwralcohol <- ggplot(data = hd, aes(sample = pwr_alcohol)) +
  stat_qq(col = "#D55E00") +
  ggtitle("Quantile Plot - alcohol (power = 0.4)") +
  theme_minimal() +
  theme(plot.background = element_rect(colour = "black", fill=NA, size=1))

qplt_pwralcohol
```

\subsection{Nine - Logistic Regression with Transformed Variables}

```{r}
logit_model_transformed <- glm(formula = chd ~ adiposity + typea + age +
                                 pwr_sbp + pwr_tobacco + pwr_ldl +
                                 pwr_obesity + pwr_alcohol, 
                               family = binomial(link = "logit"), data = hd)

summary(logit_model_transformed)
```

\newpage

\section{Problem 2 - Reporting}

\subsection{One - Univariate Table}

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.pos='H'}
library(psych)
library(kableExtra)
library(dplyr)

# select only the needed columns and gather statistics
df <- select(hd, -c(names, famhist, chd, pwr_alcohol, pwr_ldl, pwr_obesity, pwr_sbp, pwr_tobacco))
tbl <- describe(df, skew = TRUE)
# gather only mean, median and skewness
tbl <- tbl[, c(3, 5, 11)]
# print tbl
kable(tbl, caption = "Univariate Statistics for Heart Disease Data", 
      linesep = "\\addlinespace", digits = 3, 
      booktabs = T, format = 'pandoc')
```

\subsection{Two - Histogram and Quantile Plot of Tobacco}

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.show="hold", out.width="50%"}
plt_tobacco
qplt_tobacco
```

\subsection{Three - Histogram and Quantile Plot of Alcohol}

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.show="hold", out.width="50%"}
plt_alcohol
qplt_alcohol
```

\subsection{Four - Univariate Table for Power Transformation}

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.pos='H'}
library(psych)
library(kableExtra)
library(dplyr)

# select only the needed columns and gather statistics
df <- select(hd, c(pwr_alcohol, pwr_ldl, pwr_obesity, pwr_sbp, pwr_tobacco))
tbl <- describe(df, skew = TRUE)
# gather only mean, median and skewness
tbl <- tbl[, c(3, 5, 11)]
# print tbl
kable(tbl, caption = "Univariate Statistics for Heart Disease Data", 
      linesep = "\\addlinespace", digits = 3, 
      booktabs = T, format = 'pandoc')
```

\subsection{Five - Histogram and Quantile Plot of Tobacco (Power Transformation)}

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.show="hold", out.width="50%"}
plt_pwrtobacco
qplt_pwrtobacco
```

\subsection{Six - Histogram and Quantile Plot of Alcohol (Power Transformation)}

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.show="hold", out.width="50%"}
plt_pwralcohol
qplt_pwralcohol
```

\subsection{Seven - Confidence Interval (Tobacco/Alcohol)}



\subsection{Eight - Confidence Interval (Alcohol)}



\subsection{Nine - Model Performance (c-statistics)}



\section{Textbook Questions}

\subsection{Question 2}
\begin{figure}[H]
\fbox{\includegraphics[width=11cm, height=22cm]{q2.png}}
\end{figure}

\subsection{Question 6}

\begin{enumerate}[label = (\alph*)]

\item Givens: $\hat{\beta_0} = -6$, $\hat{\beta_1} = 0.05$, $X_1 = 40$ hrs, $X_2 = 3.5$
\begin{gather*}
\hat{p}(X) = \frac{e^{-6 + 0.05X_1 + X_2}}{1 + e^{-6 + 0.05X_1 + X_2}} \\
= \frac{e^{-6 + 0.05(40) + X_2}}{1 + e^{-6 + 0.05(40) + X_2}} \\
= \frac{e^{-0.5}}{1 + e^{-0.5}} \\
= 0.37754
\end{gather*}

\item The student in part (a) needs to study 50 hours in order to have a 50\% chance of getting an A in the class. Givens: $\hat{p}(X) = 0.5$, $X_2 = 3.5$

\begin{gather*}
\hat{p}(X) = \frac{e^{-6 + 0.05X_1 + X_2}}{1 + e^{-6 + 0.05X_1 + X_2}} \\
0.5 = \frac{e^{-6 + 0.05X_1 + 3.5}}{1 + e^{-6 + 0.05X_1 + 3.5}} \\
0.5(1 + e^{-6 + 0.05X_1 + 3.5}) = e^{-6 + 0.05X_1 + 3.5} \\
0.5 + 0.5e^{-2.5 + 0.05X_1} = e^{-2.5 + 0.05X_1} \\
0.5 = 0.5e^{-2.5 + 0.05X_1} \\
1 = e^{-2.5 + 0.05X_1} \\
log(1) = log(e^{-2.5 + 0.05X_1}) \\
0 = -2.5 + 0.05X_1 \\
2.5 = 0.05X_1 \\
X_1 = 50
\end{gather*}


\end{enumerate}

\subsection{Question 7}

\begin{figure}[H]
\fbox{\includegraphics[width=11cm, height=22cm]{q7.png}}
\end{figure}

\subsection{Question 8}

Since we are unsure what the true test data error rate is for the K Nearest neighbor we should not use this method. It could be that the error rate for the test data was 30% while the training data is 6%, but without knowing for sure it would be unwise to use this method. Additionally, if you multiply the average error rate by 2, then you get 36% which is greater than the 30% for logistic regression. With this being said it would be best to use logistic regression.

\subsection{Question 10}



\subsection{Question 12}


