---
title: "STA5206 Final Project"
author: "Kyle Scott"
date: "11/18/2021"
documentclass: report
geometry: margin=2cm
header-includes:
    - \usepackage{setspace}\doublespacing

output:
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question One

### We seek to find out whether a liquid diet has successfully lowered the weight of 12 adult subjects. For the purpose of this study, we will use the Wilcoxon Ranked Sum Test to test the following hypothesis:

The Null Hypothesis -> $H_0 : \mu_{t1} = \mu_{t2}$ 

The Alternative Hypothesis -> $H_1 : \mu_{t1} > \mu_{t2}$ 

* t1 represents the weights of the 12 subjects before the liquid diet and t2 represents the weights of the 12 subjects after the liquid diet has been completed.

$~$

```{r, warning=FALSE}
# create the columns for time 1 and time 2
t1 <- c(186, 171, 177, 168, 191, 172, 177, 191, 170, 171, 188, 187)
t2 <- c(188, 177, 176, 169, 196, 172, 165, 190, 166, 180, 181, 172)
# form a dataframe for our data
df <- data.frame(t1, t2)
# we must use a paired test here
  # t1 and t2 have data from the same subject -> paired
wilcox.test(df$t1, df$t2, paired = TRUE, alternative = "greater")
```

$~$

With a p-value of 0.3442, we can comfortably state that the null hypothesis cannot be rejected. This is analogous to say that the diet plan had no positive effect on the weight of the subjects after the diet was completed.

# Question Two

### We are asked to determine whether the yields of sweet potatoes from 4 different samples are the same. For the purpose of this problem we will use the Kruskal Wallis Test.

The Null Hypothesis: The is no difference between the distribution of the 4 samples

The Alternative Hypothesis: There is a significant difference between the distributions of the 4 samples

$~$

```{r}
# create the column for yield
yield <- c(8.3, 9.4, 9.1, 9.1, 9.0, 8.9, 8.9, 9.1, 9.0, 8.1, 8.2, 8.8,
           8.4, 8.3, 10.1, 10.0, 9.6, 9.3, 9.8, 9.5, 9.4, 7.8, 8.2,
           8.1, 7.9, 7.7, 8.0, 8.1)
# create the column for group
group <- c("a", "a", "a", "a", "a", "a", "a", "b", "b", "b", 
           "b", "b", "b", "b", "c", "c", "c", "c", "c", "c", "c", 
           "d", "d", "d", "d", "d", "d", "d")
# create the dataframe
sp_df <- data.frame(yield, group)
```

$~$

```{r, warning=FALSE, message=FALSE}
# we should get an idea of what our data looks like first
library("ggpubr")
library("ggthemes")
# use ggplot to create a jitter of each group and connect them by MSE
ggline(sp_df, x = "group", y = "yield", 
       color = "#99D8C9",
       add = c("mean_se", "jitter"), 
       order = c("a", "b", "c", "d"),
       ylab = "Yield", xlab = "Group", 
       title = "Group Distributions for Sweet Potato Yield", 
       ggtheme = theme_stata())
```

$~$

We can see here that our 4 samples do appear to have differences in their distributions (and their means for that matter, even though that isn't what we're testing here). It should be expected to find significance when we run our test.

$~$

```{r}
# Kruskal Wallis rank sum test using 'yield' by 'group'
kruskal.test(yield ~ group, data = sp_df)
```

$~$

As we expected based upon our visualization, we have found our p-value to be essentially zero, therefore we state that there is a significant difference between the distributions of the four samples in this test.

# Question Three

### We have a random sample of 401 single persons split by gender and asked "Would you consider marrying someone who was $25,000 or more in debt?" because it is common for people out of college to have built up a lot of debt from their studies. We are first tasked with finding (a) the expected counts for the 2x3 table of survey results.

$~$

```{r}
# create the 2x3 table as given in the question
tbl_debt <- as.table(rbind(c(125, 59, 21), c(101, 79, 16)))
dimnames(tbl_debt) <- list(Gender = c("Women", "Men"),
                      Response = c("Yes", "No", "Uncertain"))
# run a Chi-Square test on the contingency table
ChiSQ_debt <- chisq.test(tbl_debt)

# create a new table to display observed counts with expected counts in parentheses
tbl_debt2 <- tbl_debt;
tbl_debt2[] <- paste(tbl_debt,paste0("(",round(ChiSQ_debt$expected, digits = 2),")"))

# print the new table
tbl_debt2
```

$~$

### Next, we are tasked with (b) testing whether gender and the survey response are associated. We will use the Chi-Square test for independence with the following hypotheses:

The Null Hypothesis -> Gender and Survey Results are independent in this study

The Alternative Hypothesis -> Gender and Survery Results are not independent and therefore linked in some way in this study

We will test at the ${\alpha} = 0.05$ level.

$~$

```{r}
# display the Chi-Square test produced in the last section
ChiSQ_debt
```

$~$

With a p-value of 0.05172, we can conclude that at the 0.05 significance level we do not have sufficient evidence to reject the null hypothesis. This means we are unable to conclude that Gender and Survey Results are significantly linked. It is important to note that are p-value is very close to the 0.05 significance level and would be rejected at the 0.1 significance level. For the purpose of our study, however, we are unable to come to that conclusion.

# Question Four

### An experiment takes place where 5 executives are asked to rate three different methods of measuring risk on a scale of 0 to 20; the higher the number the higher the confidence in the method. These scales are to be taken as normally distributed. Given that the variance of all scores is found to be 26.257, we will now (a) find the missing values to our ANOVA table.

```{r}
# type out the data
response <- c(1.3, 4.8, 9.2, 2.5, 6.9, 14.4, 7.2, 9.1, 16.5, 6.8,
              13.2, 17.6, 12.6, 13.6, 15.5)
executive <- gl(n = 5, k = 3)
method <- gl(3, 1, length = 15)

# run aov on the data
aov_exec <- aov(response ~ method + executive)

# print the anova table
anova(aov_exec)
```

Most of our required results can be displayed here. The values 1-11 are as follows

\begin{enumerate}
\item 2
\item 4
\item 8
\item 14
\item 185.54
\item 152.24
\item 367.6
\item 92.77
\item 38.06
\item 3.73
\item 24.88
\end{enumerate}
